### 2021-10-29 text/outline/outline.tex
finish draft of part 1

Despite their promise, single-cell CRISPR screens present substantial statistical challenges. A major difficulty is that CRISPR perturbations are unobservable and assigned stochastically to cells. As a consequence, one cannot know with certainty which cells were perturbed. Instead, one must leverage an indirect, noisy proxy of perturbation presence or absence -- namely, transcribed barcode counts -- to ``guess'' which cells were perturbed. Using these imputed perturbation assignments, one can attempt to estimate the effect of the perturbation on gene expression. The standard approach, which we call ``thresholded regression'' or the ``thresholding method,'' is to assign perturbation identities to cells by simply thresholding the barcode counts.

We study estimation and inference in single-cell CRISPR screens from a statistical perspective, formulating the data generating mechanism using a new class of errors-in-variables (or measurement error) models. We assume that the response variable $y$ is a GLM of an underlying predictor variable $x^*$. We do not observe $x^*$ directly; rather, we observe a noisy version $x$ of $x^*$ that itself is a GLM of $x^*$. The goal of the analysis is to estimate the effect of $x^*$ on $y$ using the observed data $(x , y)$ only. In the context of the biological application, $x^*$, $y$, and $x$ are CRISPR perturbations, gene expressions, and barcode counts, respectively.

@@

The genomics literature has produced several applied methods for single-cell CRISPR screen analysis. In a prior work we developed SCEPTRE \cite{Barry2020}, a custom implementation of the conditional randomization test \cite{Candes2018, Liu2021} tailored to single-cell CRISPR screen data. SCEPTRE tests whether a given perturbation is associated with the change in expression of a given gene, adjusting for sources of confounding and ensuring robustness to expression model misspecification. Other applied methods for single-cell CRISPR screen analysis include MIMOSCA \cite{Dixit2016} and scMAGeCK \cite{Yang2019}. These methods, like SCEPTRE, focus on hypothesis testing (rather than estimation), but unlike SCEPTRE, they ignore the countedness of the data and are unable to handle confounders. In this work we tackle a set of analysis challenges that are complimentary to the challenges addressed by SCEPTRE; most importantly, we seek to \textit{estimate} (with confidence) the effect size of a perturbation on gene expression change, moving beyond hypothesis testing.

\subsection{Assay overview and analysis challenges}

There are several broad classes of single-cell CRISPR screen assays, each suited to answer a different set of biological questions \cite{Gasperini2019,Datlinger2021,Mimitou2019}. In this work we focus on so-called high-multiplicity of infection (MOI) single-cell CRISPR screens. We expect the ideas that we develop for this assay to apply (with some effort) to other classes of single-cell CRISPR screens as well. This section motivates high MOI single-cell screens and provides some background on the experimental protocol and relevant analysis challenges.

The human genome consists of genes, enhancers (segments of DNA that regulate the expression of one or more genes), and other genomic elements (that are not of importance to the current discussion). Genome-wide association studies (GWAS) have revealed that the majority ($>90\%$) of variants associated with diseases lie outside genes and (very likely) inside enhancers \cite{Gallagher2018}. These noncoding variants are thought to contribute to disease by modulating the expression one or more disease-relevant genes. Unfortunately, we do not know which gene (or genes) through which most noncoding variants exert their effect, limiting the interpretability of GWAS results. A central open challenge in genetics, therefore, is to link enhancers that harbor GWAS variants to the genes that they target at genome-wide scale \cite{Gasperini2020}.

The most promising biotechnology for solving this challenge are high MOI single-cell CRISPR screens. High MOI single-cell CRISPR screens combine CRISPR interference (CRISPRi) -- a variant of CRISPR that \textit{represses} a targeted region of the genome, leaving the primary DNA sequence intact  -- with single-cell sequencing. The experimental protocol, at a high level, is as follows. First, the scientist designs a library of several hundred to several thousand CRISPRi perturbations, each targeting a candidate enhancer for repression. Next, the scientist cultures several hundred thousand cells and delivers the CRISPRi perturbations to these cells. The perturbations assort into the cells randomly, with each cell receiving on average 10-40 distinct perturbations.

After waiting several days for CRISPRi to take effect, the scientist profiles each cell's transcriptome (i.e., its set of gene expressions) and perturbations. Finally, the scientist conducts perturbation-to-gene association analyses. Figure \ref{analysis_challenges}a depicts this process schematically, with colored bars (blue, red, and purple) representing distinct perturbations. For a given perturbation (e.g., the perturbation represented in blue), the scientist partitions the cells into two sets: those that received the perturbation (top) and those that did not (bottom). Next, for a given gene, the scientist runs a differential expression analysis across the two groups of cells, producing an estimate for the change in expression in response to the perturbation. If the estimated change in expression is large, the scientist can conclude that the enhancer targeted by the perturbation exerts a strong regulatory effect on the gene. This procedure is repeated for a large set of preselected perturbation-gene pairs.

% the scientist carries out the standard single-cell sequencing procedure to assay each cell's transcriptome -- -- and

% Discuss bardodes and gRNAs in the subsequent paragraph.
% \textbf{P3}: Single-cell CRISPR screens pose several core analysis challenges. \blue{Describe the analysis challenges here: (i) unobserved perturbation; (ii) existence of background reads; (iii) highly discrete count data; (iv) nuisance variables}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=1\linewidth]{../../figures/analysis_challenges/plot.pdf}
	\caption{\textbf{Experimental design and analysis challenges}: \textbf{a,} Experimental design. For a given perturbation (e.g., the perturbation represented in yellow), we partition the cells into two groups: those that received the perturbation, and those that did not receive the perturbation. For a given gene, we conduct a differential expression analysis across the two groups of cells, yielding an estimate of the impact of the given perturbation on the given gene. \textbf{b,} DAG representing the variables in the analysis. The perturbation (unobserved) affects both gene expression and gRNA expression; technical factors (e.g., batch, sequencing depth, etc.) act as nuisance variables. The target of inference is the effect of the perturbation on gene expression (denoted with question mark). \textbf{c,} Schematic illustrating ``background reads.'' The gRNA modality has a nonzero, ``background read'' distribution even in the absence of a perturbation, complicating the assignment of perturbations to cells. \textbf{d}, Example data for a given perturbation-gene pair. Notice that (i) the perturbations are unobserved, and (ii) the gene and gRNA expression data take the form of discrete counts.}
	\label{analysis_challenges}
\end{figure}

@@

In this section we study thresholded regression from empirical and theoretical perspectives, highlighting several limitations of the method. Let $n \in \N$ be the number of cells assayed in the experiment (typically, $n \approx 100,000 - 250,000$). Consider a single perturbation and a single gene. For cell $i \in \{1, \dots, n\}$, let $m_i \in \N$ be the number of gene transcripts sequenced in the cell; let $g_i \in \N$ be the number of gRNA transcripts sequenced in the cell; let $l^m_i \in \N$ be the number of gene transcripts sequenced across \textit{all} genes (the library size) in the cell; and finally, let $z_i \in \R^{d-1}$ be the cell-specific vector of technical factors (e.g., sequencing batch, percent mitochondrial reads, etc.). The thresholding method is defined as follows:
\begin{itemize}
\item[1.] For a given threshold $c \in \N$, let the imputed perturbation assignment $\hat{p}_i \in \{0, 1\}$ be $$\begin{cases} \hat{p}_i = 0 \textrm{ if } g_i \geq c, \\ \hat{p}_i = 1 \textrm{ if } g_i < c. \end{cases}$$
\item[2.] Assume that $m_i$ is related to $\hat{p}_i, l_i,$ and $z_i$ through the following GLM:
\begin{equation*}
\begin{cases}
m_i | (\hat{p}_i, z_i, l_i) \sim \textrm{NB}_\theta(\mu_i), \\
\log(\mu_i) = \beta^m_0 + \beta^m_1 \hat{p}_i + \gamma^T_m z_i + \log\left( l_i^m\right),
\end{cases}
\end{equation*}
where (i) $NB_\theta(\mu_i)$ is a negative binomial distribution with mean $\mu_i$ and known size parameter $\theta$; (ii) $\beta^m_0 \in \R, \beta^m_1 \in \R,$ and $\gamma \in  \R^{d-1}$ are unknown parameters; and (iii) $\log(l_i^m)$ is an offset term. Fit a GLM to obtain estimates of the parameters.

@@

The biological interpretation for the target of inference $\beta^m_1$ as follows:  $\beta^m_1$ is the log-transformed fold change in gene expression in response to the perturbation, removing the effect of the technical factors. Fold change (obtained by exponentiating $\beta^m_1$) is the ratio of the mean expression in cells that were perturbed to the mean expression in cells that were not perturbed. $\exp(\beta^m_1) = 1$ indicates no change in mean expression, while $\exp(\beta^m_1) > 1$ and $\exp(\beta^m_1) < 1$ indicate an increase and decrease in expression in response to the perturbation, respectively (accounting for technical factors).

\subsection{Empirical analysis}

@@

We reasoned that thresholded regression underestimated effect sizes on the positive control pairs, especially for threshold = 1 (an example of \textit{attenuation bias}). For a given perturbation, the vast majority ($>99\%$) of cells are unperturbed. This imbalance leads to an asymmetry: misclassifying \textit{unperturbed} cells as \textit{perturbed} is intuitively ``worse'' than misclassifying \textit{perturbed} cells as \textit{unperturbed}. Misclassified \textit{unperturbed} cells contaminate the set of truly perturbed cells, leading to attenuation bias; by contrast, misclassified \textit{perturbed} cells are swamped in number and ``neutralized'' by the truly \textit{unperturbed} cells. Setting the threshold to a large number reduces the unperturbed-to-perturbed misclassification rate, decreasing bias.

We hypothesized, however, that the reduction in bias conferred by selecting a large threshold incurs the cost of increasing the variance of the estimator. To investigate, we compared $p$-values and confidence intervals produced by threshold = 5 and threshold = 20 for the target of inference $\beta^m_1$. We found that threshold = 5 yielded smaller (i.e., more significant) $p$-values and narrower confidence intervals than did threshold = 20 (Figure \ref{thresholding_empirical}c-d). We concluded that thresholded regression has a bias-variance tradeoff: as the threshold increases, bias decreases and variance increases.

@@

\subsection{Theoretical analysis}\label{sec:thresholding_theory}

We study the thresholding method from a theoretical perspective in an idealized Gaussian setting. Suppose that we observe data $(g_1, m_1), \dots, (g_n, m_n)$ from the following model:

@@

\begin{proposition}\label{prop:convergence}
 The almost sure limit (as $n \to \infty$) of $\hat{\beta}^m_1$ is
\begin{equation}\label{thresh_est_intercepts}
\hat{\beta}^m_1 \xrightarrow{a.s.} \beta^m_1 \left(\frac{ \pi( \omega - \E[ \hat{p}_i ])}{ \E[\hat{p}_i] (1 - \E[\hat{p}_i])}\right),
\end{equation}
where
$$\begin{cases}
\E[\hat{p}_i] = \zeta(1-\pi) + \omega\pi , \\
\omega = \Phi\left(\beta_1^g + \beta_0^g -c \right) ,\\ \zeta = \Phi\left( \beta^g_0 - c \right). \\
\end{cases}$$
\end{proposition}
Let $\gamma: \R^4 \to \R$ be defined by
$$ \gamma(\beta^g_1, \pi, c, \beta^g_0) = \frac{\pi (\omega - \E[\hat{p}_i])}{\E[\hat{p}_i] (1 - \E[\hat{p}_i]) }.$$ We call $\gamma$ the ``attenuation function.'' Observe that \begin{itemize}
\item[i.] $\gamma$ does not depend on $\beta^m_1$ or $\beta^m_0$, and
\item[ii.] $\hat{\beta}^m_1 \xrightarrow{a.s.} [\gamma(\beta_0^g, \beta_1^g, c, \pi)] \beta^m_1.$
\end{itemize}
Let $b: \R^4 \to \R$ be the asymptotic \textit{relative} bias of $\hat{\beta}^m_1$:
\begin{multline*}
b(\beta^g_1, \pi, c, \beta^g_0) = \left(\frac{1}{\beta^m_1}\right) \lim_{n\to\infty} \left(\beta^m_1 - \E[\hat{\beta}^m_1]\right) = \left( \frac{1}{\beta^m_1}\right)\left(\beta^m_1 - \E \left(\lim_{a.s.} \hat{\beta}^m_1 \right)\right) \\ = \frac{1}{\beta^m_1} \left(\beta^m_1 - \gamma(\beta^g_1, \pi, c, \beta^g_0)\beta^m_1\right) = 1 - \gamma(\beta^g_1, \pi, c, \beta^g_0),
\end{multline*}
where $\lim_{a.s.}$ denotes a.s.\ convergence. The asymptotic relative bias vanishes when the attenuation function equals $1$. \textcolor{red}{(HOW TO MAKE THE ABOVE RIGOROUS?)}

@@

To investigate the basic question of ``What is a good threshold selection strategy?'', we study the relationship between the asymptotic relative bias $b$ of $\hat{\beta}^m_1$ and the selected threshold $c$. For simplicity, we set the perturbation probability $\pi$ to $1/2$. Let $c_\textrm{bayes} \in \R$ be the Bayes-optimal decision boundary for classifying cells as perturbed or unperturbed, i.e. $$c_\textrm{bayes} = \argmin_{c \in \R} \P(\hat{p}_i \neq p_i).$$ Simple algebra shows that $c_\textrm{bayes} = \beta_0^g + (1/2) \beta^g_1.$ Below, we give several results for the asymptotic relative bias $b$ of $\hat{\beta}^m_1$. We refer throughout to Figure \ref{thresholding_theoretical}a, which displays plots of asymptotic relative bias versus threshold for different values of $\beta^g_1$. We sometimes refer to ``asymptotic relative bias'' using the shortened term ``bias'' for succinctness.

@@

\begin{center}
\textbf{Generalizing to $\pi \in [0,1/2]$ (Panel b)}
\end{center}
We generalize the expression for bias when the threshold is large to arbitrary $\pi \in [0,1/2]$:
\begin{proposition}\label{prop:c_limit}
Assume without loss of generality that $\beta^g_1 > 0$. As the threshold $c$ tends to infinity, the asymptotic relative bias $b$ tends to $\pi,$ i.e.
$$ \lim_{ c \to \infty } b(\beta^g_1, \pi, c, \beta^g_0) = \pi.$$
\end{proposition}
In other words, if the perturbation probability is $\pi$, and if we set the threshold to a large number, then the asymptotic relative bias is $\pi$ (Figure \ref{thresholding_theoretical}b). We can understand this result intuitively by considering an extreme example: when $\pi$ is very small (e.g., $\pi = 0.01$), almost all cells are unperturbed. Therefore, in selecting a large threshold, we correctly classify nearly all unperturbed cells as unperturbed; on the other hand, the \textit{perturbed} cells that we misclassify as \textit{unperturbed} are swamped in number by the truly unperturbed cells, resulting in a small bias.

>>>>

Despite their promise, single-cell CRISPR screens present substantial statistical challenges. A major difficulty is that CRISPR perturbations are unobservable and assigned stochastically to cells. As a consequence, one cannot know with certainty which cells were perturbed. Instead, one must leverage an indirect, noisy proxy of perturbation presence or absence -- namely, transcribed guide RNA counts -- to ``guess'' which cells were perturbed. Using these imputed perturbation assignments, one can attempt to estimate the effect of the perturbation on gene expression. The standard approach, which we call ``thresholded regression'' or the ``thresholding method,'' is to assign perturbation identities to cells by simply thresholding the guide RNA counts.

We study estimation and inference in single-cell CRISPR screens from a statistical perspective, formulating the data generating mechanism using a new class of errors-in-variables (or measurement error) models. We assume that the response variable $y$ is a GLM of an underlying predictor variable $x^*$. We do not observe $x^*$ directly; rather, we observe a noisy version $x$ of $x^*$ that itself is a GLM of $x^*$. The goal of the analysis is to estimate the effect of $x^*$ on $y$ using the observed data $(x , y)$ only. In the context of the biological application, $x^*$, $y$, and $x$ are CRISPR perturbations, gene expressions, and guide RNA counts, respectively.

@@

The genomics literature has produced several applied methods for single-cell CRISPR screen analysis. In a prior work we developed SCEPTRE \cite{Barry2020}, a custom implementation of the conditional randomization test \cite{Candes2018, Liu2021} tailored to single-cell CRISPR screen data. SCEPTRE tests whether a given perturbation is associated with the change in expression of a given gene, adjusting for sources of confounding and ensuring robustness to expression model misspecification. Other applied methods for single-cell CRISPR screen analysis include MIMOSCA \cite{Dixit2016} and scMAGeCK \cite{Yang2019}. These methods, like SCEPTRE, focus on hypothesis testing (rather than estimation), but unlike SCEPTRE, they ignore the countedness of the data and are unable to handle confounders. In this work we tackle a set of analysis challenges that are complimentary to the challenges addressed by SCEPTRE. Most importantly, we seek to \textit{estimate} (with confidence) the effect size of a perturbation on gene expression change, moving beyond hypothesis testing.

\subsection{Assay overview}

\begin{figure}
	\centering
	\includegraphics[width=1\linewidth]{../../figures/analysis_challenges/plot.pdf}
	\caption{\textbf{Experimental design and analysis challenges}: \textbf{a,} Experimental design. For a given perturbation (e.g., the perturbation indicated in blue), we partition the cells into two groups: perturbed and unperturbed. Next, for a given gene, we conduct a differential expression analysis across the two groups, yielding an estimate of the impact of the given perturbation on the given gene. \textbf{b,} DAG representing all variables in the system. The perturbation (latent) impacts both gene expression and gRNA expression; technical factors act as nuisance variables, also impacting gene and gRNA expression. The target of estimation is the effect of the perturbation on gene expression. \textbf{c,} Schematic illustrating the ``background read'' phenomenon. Due to errors in the sequencing process, unperturbed cells exhibit a nonzero gRNA count distribution (bottom). The target of estimation is the change in mean gene expression in response to the perturbation (top). \textbf{d}, Example data on four cells for a given perturbation-gene pair. Note that (i) the perturbation is unobserved, and (ii) the gene and gRNA data are discrete counts.}
	\label{analysis_challenges}
\end{figure}

There are several broad classes of single-cell CRISPR screen assays, each suited to answer a different set of biological questions \cite{Gasperini2019,Datlinger2021,Mimitou2019}. In this work we focus on so-called high-multiplicity of infection (MOI) single-cell CRISPR screens. We expect the ideas that we develop for this assay to apply (with some effort) to other classes of single-cell CRISPR screens as well. In this section we motivate high MOI single-cell screens, overview the experimental protocol, and present relevant analysis challenges.

The human genome consists of genes, enhancers (segments of DNA that regulate the expression of one or more genes), and other genomic elements (that are not of importance to the current discussion). Genome-wide association studies (GWAS) have revealed that the majority ($>90\%$) of variants associated with diseases lie outside genes and (very likely) inside enhancers \cite{Gallagher2018}. These noncoding variants are thought to contribute to disease by modulating the expression one or more disease-relevant genes. We do not know the gene (or genes) through which most noncoding variants exert their effect, limiting the interpretability of GWAS results. A central open challenge in genetics, therefore, is to link enhancers that harbor GWAS variants to the genes that they target at genome-wide scale \cite{Gasperini2020}.

The most promising biotechnology for solving this challenge are high MOI single-cell CRISPR screens. High MOI single-cell CRISPR screens combine CRISPR interference (CRISPRi) -- a version of CRISPR that \textit{represses} a targeted region of the genome, leaving the primary DNA sequence intact  -- with single-cell sequencing. The experimental protocol is as follows. First, the scientist develops a library of several hundred to several thousand CRISPRi perturbations, each designed to target a candidate enhancer for repression. The scientist then cultures tens or hundreds of thousands of cells and delivers the CRISPRi perturbations to the cells. The perturbations assort into the cells randomly, with each cell receiving on average 10-40 distinct perturbations. Conversely, a given perturbation enters fewer than 0.1-2\% of cells.

After waiting several days for CRISPRi to take effect, the scientist profiles each cell's transcriptome (i.e., its gene expressions) and the set of perturbations that it received. Finally, the scientist conducts perturbation-to-gene association analyses. Figure \ref{analysis_challenges}a depicts this process schematically, with colored bars (blue, red, and purple) representing distinct perturbations. For a given perturbation (e.g., the perturbation represented in blue), the scientist partitions the cells into two groups: those that received the perturbation (top) and those that did not (bottom). Next, for a given gene, the scientist runs a differential expression analysis across the two groups of cells, producing an estimate for the magnitude of gene expression change in response to the perturbation. If the estimated change in expression is large, the scientist can conclude that the enhancer \textit{targeted} by the perturbation exerts a strong regulatory effect on the gene. This procedure is repeated for a large set of preselected perturbation-gene pairs. % If the differential expression analyses produce $p$-values, the $p$-values are subjected to a multiple hypothesis testing correction procedure.

\subsection{Analysis challenges and data}
High MOI single-cell CRISPR screens present several statistical challenges. We take a parametric approach in this work, and so we highlight challenges that are especially relevant to modeling the data generating process. Throughout, we consider a single perturbation-gene pair. First, the ``treatment'' variable -- i.e., the presence or absence of a perturbation -- is unobservable. Perturbed cells transcribe molecules called  \textit{guide RNAs} (or \textit{gRNAs}) that serve as indirect signals of perturbation presence. We can leverage the gRNA data to infer whether a given cell has been perturbed. Using these imputed perturbation assignments, we can estimate the effect of the perturbation on gene expression (Figure \ref{analysis_challenges}b). Second, ``technical factors'' -- sources of variation that are experimental rather than biological in origin -- impact both gene expression and gRNA expression (Figure \ref{analysis_challenges}b).  We must account for these technical factors in constructing models for gene and gRNA expression. Third, the gene and gRNA data are sparse, discrete counts. Therefore, classical statistical approaches that assume Gaussianity or homoscedasticity are inapplicable. Finally, and most subtly, sequenced gRNAs sometimes are mapped to unperturbed cells. This phenomenon, which we call the ``background read'' phenomenon, results from errors in the sequencing process \cite{Replogle2020}. The marginal distribution of the gRNA count data is best conceptualized as a mixture model (Figure \ref{analysis_challenges}c; Gaussian distributions used for illustration purposes only). Unperturbed and perturbed cells both exhibit nonzero gRNA count distributions, but the gRNA count distribution for perturbed cells is greater. Figure \ref{analysis_challenges}d shows example data for four (of possibly tens or hundreds of thousands of) cells. The analysis objective is to leverage the observed data -- i.e., the gene expressions and gRNA counts -- to estimate the effect of the (latent) perturbation on gene expression, accounting for the technical factors.

In this work we analyze two large-scale, high MOI, single-cell CRISPRi screen datasets published by Gasperini et al.\ and Xie et al. Gasperini (resp., Xie) targeted approximately 6,000 (resp., 500) candidate enhancers in a population of approximately 200,000 (resp., 100,000) cells. Gasperini additionally designed 381 positive control, gene-targeting perturbations and 50 non-targeting, negative control perturbations to assess method sensitivity and specificity.

@@

The method that Gasperini et al.\ used in their original analysis is the so-called ``thresholding method,'' or ``thresholded regression.'' In this section we study the thresholding method from empirical and theoretical perspectives, uncovering several basic limitations of the approach. Let $n \in \N$ be the number of cells assayed in the experiment. Consider a single perturbation and a single gene. For cell $i \in \{1, \dots, n\}$, let $m_i \in \N$ be the number of gene transcripts sequenced; let $g_i \in \N$ be the number of gRNA transcripts sequenced; let $l^m_i \in \N$ be the number of gene transcripts sequenced across \textit{all} genes (the library size); and finally, let $z_i \in \R^{d-1}$ be the cell-specific technical factors (e.g., sequencing batch, percent mitochondrial reads, etc.) The letters ``m,'' ``g'', and ``l'' stand for ``mRNA,'' ``gRNA,'' and ``library,'' respectively. The thresholding method is defined as follows:
\begin{itemize}
\item[1.] For a given threshold $c \in \N$, let the imputed perturbation assignment $\hat{p}_i \in \{0, 1\}$ be $$\begin{cases} \hat{p}_i = 0 \textrm{ if } g_i \geq c, \\ \hat{p}_i = 1 \textrm{ if } g_i < c. \end{cases}$$
\item[2.] Assume that $m_i$ is related to $\hat{p}_i, l_i,$ and $z_i$ through the following GLM:
\begin{equation*}
\begin{cases}
m_i | (\hat{p}_i, z_i, l_i) \sim \textrm{NB}_\theta(\mu_i), \\
\log(\mu_i) = \beta^m_0 + \beta^m_1 \hat{p}_i + \gamma^T_m z_i + \log\left( l_i^m\right),
\end{cases}
\end{equation*}
where (i) $NB_\theta(\mu_i)$ is a negative binomial distribution with mean $\mu_i$ and known size parameter $\theta$; (ii) $\beta^m_0 \in \R, \beta^m_1 \in \R,$ and $\gamma \in  \R^{d-1}$ are unknown parameters; and (iii) $\log(l_i^m)$ is an offset term. Fit a GLM to obtain estimates of the parameters.

@@

The biological interpretation for the target of inference $\beta^m_1$ as follows:  $\beta^m_1$ is the log-transformed fold change in gene expression in response to the perturbation, removing the effect of the technical factors. Fold change (obtained by exponentiating $\beta^m_1$) is the ratio of the mean gene expression in cells that were perturbed to the mean gene expression in cells that were not perturbed. $\exp(\beta^m_1) = 1$ indicates no change in mean expression, while $\exp(\beta^m_1) > 1$ and $\exp(\beta^m_1) < 1$ indicate an increase and decrease in expression in response to the perturbation, respectively (accounting for technical factors).

\subsection{Empirical challenges of thresholding method}\label{sec:thresholding_empirical}

@@

We reasoned that thresholded regression systematically underestimated effect sizes on the positive control pairs, especially for small thresholds (an example of \textit{attenuation bias}). For a given perturbation, the vast majority ($>98\%$) of cells are unperturbed. This imbalance leads to an asymmetry: misclassifying \textit{unperturbed} cells as \textit{perturbed} is intuitively ``worse'' than misclassifying \textit{perturbed} cells as \textit{unperturbed}. Misclassified unperturbed cells contaminate the set of truly perturbed cells, leading to attenuation bias; by contrast, misclassified perturbed cells are swamped in number and ``neutralized'' by the truly unperturbed cells. Setting the threshold to a large number reduces the unperturbed-to-perturbed misclassification rate, decreasing bias.

We hypothesized, however, that the reduction in bias conferred by selecting a large threshold comes with the cost of increasing the variance of the estimator. To investigate, we compared $p$-values and confidence intervals produced by threshold = 5 and threshold = 20 for the target of inference $\beta^m_1$. We found that threshold = 5 yielded smaller (i.e., more significant) $p$-values and narrower confidence intervals than did threshold = 20 (Figure \ref{thresholding_empirical}c-d). We concluded that the threshold controls a bias-variance tradeoff: as the threshold increases, bias decreases and variance increases.

@@

\subsection{Theoretical challenges of thresholding method}\label{sec:thresholding_theory}

Next, we study the thresholding method from a theoretical perspective, recovering the attenuation bias and bias-variance tradeoff effects established in the previous section, as well as several other interesting phenomena.  We work in an idealized Gaussian setting. Suppose that we observe gRNA and gene expression data $\{(g_1, m_1), \dots, (g_n, m_n)\}$ on $n \in \N$ cells from the following model:

@@

\begin{proposition}\label{prop:convergence}
 The almost sure limit (as $n \to \infty$) of $\hat{\beta}^m_1$ is
\begin{equation}\label{thresh_est_intercepts}
\hat{\beta}^m_1 \xrightarrow{a.s.} \beta^m_1 \left(\frac{ \pi( \omega - \E[ \hat{p}_i ])}{ \E[\hat{p}_i] (1 - \E[\hat{p}_i])}\right),
\end{equation}
where
$$\begin{cases}
\E[\hat{p}_i] = \zeta(1-\pi) + \omega\pi , \\
\omega = \Phi\left(\beta_1^g + \beta_0^g -c \right) ,\\ \zeta = \Phi\left( \beta^g_0 - c \right). \\
\end{cases}$$
\end{proposition}
Let $\gamma: \R^4 \to \R$ be defined by
$$ \gamma(\beta^g_1, \pi, c, \beta^g_0) = \frac{\pi (\omega - \E[\hat{p}_i])}{\E[\hat{p}_i] (1 - \E[\hat{p}_i]) }.$$ We call $\gamma$ the ``attenuation function.'' Observe that \begin{itemize}
\item[i.] $\gamma$ does not depend on $\beta^m_1$ or $\beta^m_0$, and
\item[ii.] $\hat{\beta}^m_1 \xrightarrow{a.s.} [\gamma(\beta_0^g, \beta_1^g, c, \pi)] \beta^m_1.$
\end{itemize}
Let $b: \R^4 \to \R$ be the asymptotic relative bias of $\hat{\beta}^m_1$:
\begin{multline*}
b(\beta^g_1, \pi, c, \beta^g_0) = \left(\frac{1}{\beta^m_1}\right) \lim_{n\to\infty} \left(\beta^m_1 - \E[\hat{\beta}^m_1]\right) = \left( \frac{1}{\beta^m_1}\right)\left(\beta^m_1 - \E \left(\lim_{a.s.} \hat{\beta}^m_1 \right)\right) \\ = \frac{1}{\beta^m_1} \left(\beta^m_1 - \gamma(\beta^g_1, \pi, c, \beta^g_0)\beta^m_1\right) = 1 - \gamma(\beta^g_1, \pi, c, \beta^g_0),
\end{multline*}
where $\lim_{a.s.}$ denotes a.s.\ convergence. The asymptotic relative bias vanishes when the attenuation function equals $1$. \textcolor{red}{HOW TO MAKE THE ABOVE RIGOROUS?}

@@

To investigate the basic question of ``What is a good threshold selection strategy?'', we study the relationship between the asymptotic relative bias $b$ of $\hat{\beta}^m_1$ and the selected threshold $c$. For simplicity, we begin by setting the perturbation probability $\pi$ to $1/2$. Let $c_\textrm{bayes} \in \R$ be the Bayes-optimal decision boundary for classifying cells as perturbed or unperturbed, i.e. $$c_\textrm{bayes} = \argmin_{c \in \R} \P(\hat{p}_i \neq p_i).$$ Simple algebra shows that $c_\textrm{bayes} = \beta_0^g + (1/2) \beta^g_1.$ Below, we give several results for the asymptotic relative bias $b$ of $\hat{\beta}^m_1$. We refer throughout to Figure \ref{thresholding_theoretical}a, which displays plots of asymptotic relative bias versus threshold for different values of $\beta^g_1$. We sometimes refer to ``asymptotic relative bias'' using the shortened term ``bias'' for succinctness. \textcolor{red}{SHOULD WE MOVE PROP 4-6 	TO THE APPENDIX? THESE PROPS ARE LESS IMPORTANT THAN THE OTHERS AND CONTRIBUTE SOMEWHAT LESS TO THE NARRATIVE.}

@@

\begin{center}
\textbf{Generalizing to $\pi \in [0,1/2]$ (Panel b)}
\end{center}
We generalize the expression for bias when the threshold is large to arbitrary $\pi \in [0,1/2]$:
\begin{proposition}\label{prop:c_limit}
Assume without loss of generality that $\beta^g_1 > 0$. As the threshold $c$ tends to infinity, the asymptotic relative bias $b$ tends to $\pi,$ i.e.
$$ \lim_{ c \to \infty } b(\beta^g_1, \pi, c, \beta^g_0) = \pi.$$
\end{proposition}
In other words, if the perturbation probability is $\pi$, and if we set the threshold to a large number, then the asymptotic relative bias is $\pi$ (Figure \ref{thresholding_theoretical}b). We can understand this result intuitively by considering an extreme example: when $\pi$ is very small (e.g., $\pi = 0.01$), most cells are unperturbed. Therefore, as discussed in Section \ref{sec:thresholding_empirical}, selecting a large threshold minimizes the unperturbed-to-perturbed misclassification rate, reducing bias. % Therefore, in selecting a large threshold, we correctly classify nearly all unperturbed cells as unperturbed; on the other hand, the \textit{perturbed} cells that we misclassify as \textit{unperturbed} are swamped in number by the truly unperturbed cells, resulting in a small bias.
